---
title: "Covid91 vaccine study Final2024"
author: "Mairani G"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 Vaccine is not effective in males.

$H_A$: The Covid921 vaccine is effective in males.

### Methods

The treatment is categorical and sex is also categorical. I used the cat~cat helper file.  

### Descriptive Results

#### Graphical Results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~infected + treatment,data=males)
barchartGC(~infected + treatment,data=males, type="percent")

```

The graph shows that males who received the vaccine are less likely to get covid. People who received the placebo showed to have higher rates of covid. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```
The table shows that 73.8% of males who received the placebo became infected with covid. This is higher than the 26.2% of males that recieved the vaccine and became infected with covid.  

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher's exact test shows that the p-value is 0.00000000000000022. This means the null hypothesis of the Covid921 Vaccine not being effective in males is rejected because it is a smaller value than 0.05. 
The chi and fisher tests also show this p-value.

The odds ratio was 2.815604, showing a positive relationship between males and the effectiveness of receiving the covid921 vaccine. (This is because an odds ratio greater than 1 indicates an increased odds of the effective vaccine outcome)  


## Females

$H_0$: The Covid921 Vaccine is not effective in females.

$H_A$: The Covid921 vaccine is effective in females.

### Methods

I used the Cat~Cat helper file because both variables were categorical. Treatment is categorical and sex (female) is also categorical. 

### Descriptive Results

#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")
```

The graph shows that there is an existing difference between females that got the Covid vaccine and those that got the placebo. Those that received the placebo had a higher rate of contracting covid19.

#### Numerical

```{r}
table2 <- xtabs(~treatment + infected, data=females)
rowPerc(table2)
colPerc(table2)
```

The tables shows that among the the females that received the vaccine, 50% of them were OK and didn't get Covid19. The 65% of them that received the placebo became infected with Covid19.   

(But there was nearly no difference in people that were OK overall: placebo was 49.94 OK and vaccine was 50.06 OK)  

#### Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The chi-squared test shows a p-value of 0.00000001914, smaller than 0.05, meaning we reject the null hypothesis that the Covid921 Vaccine is not effective in females. 

The Fisher's Exact Test has a p-value of 0.00000001246, confirming the null hypothesis is rejected.

The Odds Ratio is 1.856237, this is bigger than 1 so the alternative hypothesis (The Covid921 vaccine is effective in females) is true.  


## LGBTQ
$H_0$: The Covid921 Vaccine is not effective on the LGBTQ community.

$H_A$: The Covid921 vaccine is effective on the LGBTQ community

### Methods
The Cat~Cat helper file was used because the treatment is categorical and whether or not a person is apart of the LGBTQ community is also categorical. 

### Descriptive Results

#### Graphical Results 
```{r}
LGBTQ=filter(FinalData,FinalData$LGBTQ=="gay")
barchartGC(~infected + treatment, data=LGBTQ)
barchartGC(~infected +treatment, data=LGBTQ,type="percent")
```

#### Numerical

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

The Numerical chart shows that in those the received the vaccine, there was a higher rate of infections. The numbers were higher than in the people that got the placebo.(the infection rate in the vaccine group is 4 times higher than in the placebo group). 

(For the OK group, the placebo group had a higher infection rate.)  

#### Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Inferential results give us the chi-squared test that shows a p-value of 2.266e-06 (0.000002266), this is less than 0.05 so the null hypothesis that "The Covid921 Vaccine is not effective on the LGBTQ community" is rejected.   

Fisher's Exact Tests p-value is 7.269e-07(0.0000007269), confirming the null hypothesis is rejected and proving the alternative hypothesis.

The Odds Ratio is 4.27359, this is bigger than 1 so the alternative hypothesis (The Covid921 Vaccine is not effective on the LGBTQ community) is true.   

## Drug Users

$H_0$: The Covid921 Vaccine is not effective on drug users.   

$H_A$: The Covid921 vaccine is effective on drug users. 

### Methods

Drug use is a categorical variable, (yes or no to frug use), so I used the cat~cat helper file. 

### Descriptive results

#### Graphical results 
```{r}
Druggies=filter(FinalData,FinalData$DrugUser=="yes")
barchartGC(~infected + treatment, data=Druggies)
barchartGC(~infected +treatment, data=Druggies,type="percent")
```
 
The Graph shows that there is a higher infection rate among drug users that got the vaccine. They are significantly more infected by covid19 than those that received the placebo.

The infection rate of those who got the vaccine is around 90% where around 10% of the placebo recipients got infected.

#### Numerical

```{r}
table4 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table4)
colPerc(table4)
```

The Numerical results show that 91.67 of drug users that received the vaccine became infected. Only 8.33 of drug users that received the placebo got infected. 

Those who received the placebo were also OK at a higher rate. 

#### Inferential results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The inferential results Chi-squared test shows a p-value of 1.414e-10 (0.0000000001414), this is smaller than 0.05 so the null hypothesis is rejected. (that The Covid921 Vaccine is not effective on drug users.)

The Fisher's Exact Test p-value is 3.261e-12(0.000000000003261), which also rejects the null hypothesis and suggests that the Covid19 vaccine may be effective on drug users.

The Odds Ratio is 0.08740846 
, this is less than 1 so it suggests that drug users who received the covid vaccine have lower odds of getting infected compared to those who did not get the vaccine.  ( also rejects the null hypothesis)

#### Conclusion


# Overall Results and Conclusions