---
title: "Covid91 vaccine study Final2024"
author: "Mairani G"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 Vaccine is not effective in males.

$H_A$: The Covid921 vaccine is effective in males.

### Methods

The treatment is categorical and sex is also categorical. I used the cat~cat helper file.  

### Descriptive Results

#### Graphical Results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")

```

The graph shows that males who received the vaccine are less likely to get covid. People who received the placebo showed to have higher rates of covid. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```
The table shows that 73.8% of males who received the placebo became infected with covid. This is higher than the 26.2% of males that recieved the vaccine and became infected with covid.  

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher's exact test shows that the p-value is 0.00000000000000022. This means the null hypothesis of the Covid921 Vaccine not being effective in males is rejected because it is a smaller value than 0.05. 
The chi and fisher tests also show this p-value.

The odds ratio was 2.815604, showing a positive relationship between males and the effectiveness of receiving the covid921 vaccine. (This is because an odds ratio greater than 1 indicates an increased odds of the effective vaccine outcome)  


## Females

$H_0$: The Covid921 Vaccine is not effective in females.

$H_A$: The Covid921 vaccine is effective in females.

### Methods

I used the Cat~Cat helper file because both variables were categorical. Treatment is categorical and sex (female) is also categorical. 

### Descriptive Results

#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")
```

The graph shows that there is an existing difference between females that got the Covid vaccine and those that got the placebo. Those that received the placebo had a higher rate of contracting covid19.

#### Numerical

```{r}
table2 <- xtabs(~treatment + infected, data=females)
rowPerc(table2)
colPerc(table2)
```

The tables shows that among the the females that received the vaccine, 50% of them were OK and didn't get Covid19. The 65% of them that received the placebo became infected with Covid19.   

(But there was nearly no difference in people that were OK overall: placebo was 49.94 OK and vaccine was 50.06 OK)  

#### Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The chi-squared test shows a p-value of 0.00000001914, smaller than 0.05, meaning we reject the null hypothesis that the Covid921 Vaccine is not effective in females.  

The Fisher's Exact Test has a p-value of 0.00000001246, confirming the null hypothesis is rejected.

The Odds Ratio is 1.856237, this is bigger than 1 so the alternative hypothesis (The Covid921 vaccine is effective in females) is true.  


## LGBTQ
$H_0$: The Covid921 Vaccine is not effective on the LGBTQ community.

$H_A$: The Covid921 vaccine is effective on the LGBTQ community


## Drug Users

$H_0$: The Covid921 Vaccine is not effective on drug users.   

$H_A$: The Covid921 vaccine is effective on drug users.  
### Methods

Drug use is a categorical variable, (yes or no to frug use), so I used the cat~cat helper file. 

### Descriptive results

#### Graphical results 

#### Numerical

#### Inferential results

#### Conclusion


# Overall Results and Conclusions